### 2013-06-10 doc/dissertation/document/chapters/spn_equations/spn_equations.tex
mc background edits and spell check

For this study, the fuel assemby problem with a 1-group $SP_1$
discretization was solved using MCSA with the adjoint Monte Carlo
solver using both the collision and expected value estimators and both
the Richardson and MINRES fixed point iterations. Using the results
from the previous section, a Neumann relaxation parameter of 0.7 was
used with both estimators and fixed point iterations. For the
collision estimator, a Richardson relaxation parameter of 1.1 was used
with the Richardson iteration while it was found the expected value
estimator had the best performance when a Richardson relaxation
parameter of 1.0 was used instead. In addition, ILUT preconditioning
with a drop tolerance of \sn{1}{-5} and a fill level of 5 was used
with a reduced domain approximation fill level of 100 and a threshold
of \sn{1}{-10} as in the previous analysis. For each estimator, the
number of stochastic histories used to compute the MCSA correction was
varied from 25 to 50,000 with the number of MCSA iterations required
to converge to a tolerance of \sn{1}{-8} recorded for a single
eigenvalue iteration along with the CPU time required to achieve
convergence.

@@

%%---------------------------------------------------------------------------%%
\section{MCSA Verification}
\label{sec:spn_mcsa_verification}
Through the numerical studies presented in this chapter, we have
developed a set of preconditioning techniques that permit MCSA to be
used with the $SP_N$ form of the neutron transport equation and
applied to a difficult fuel assembly criticality problem. Along with
these preconditioning techniques, a set a varying MCSA parameters were
analyzed to find those that yielded the best performance for this
particular problem. In addition, several steps were taken to mitigate
the dense composite linear operators that arise when performing
explicit MCSA preconditioning in order to achieve convergence. Using
these results, we compare MCSA directly to conventional linear solvers
that would be typically used to solve the $SP_N$ form of the transport
equation as presented here. In particular, we will use a pair of
production Krylov solvers with which we will compare direct numerical
results in order to verify MCSA solutions in this section and
iterative and CPU timing performance in the next order to gauge
feasibility and identify areas that require further research.

@@

For every combination of energy group structure and solver, the
eigenvalue problem was converged with an eigensolver tolerance of
\sn{1}{-6} and a linear solver tolerance of \sn{1}{-8} in a serial
calculation. Table~\ref{tab:serial_ev_results} gives the eigenvalue
computed by each solver for each problem as well as the number of
eigenvalue iterations required to converge. For every problem, each
solver converged to the same eigenvalue within the tolerance of the
eigensolver in the same number of eigenvalue iterations.
\begin{table}[h!]
  \begin{center}
    \begin{tabular}{lccc}\hline\hline
      \multicolumn{1}{c}{\textbf{Solver}} & 
      \multicolumn{1}{c}{\textbf{Number of Groups}} & 
      \multicolumn{1}{c}{\textbf{Iterations}} &
      \multicolumn{1}{c}{\textbf{Eigenvalue}} \\
      \hline
      BiCGStab-ILUT & 1 & 25 & 1.1059870 \\
      GMRES-ILUT & 1 & 25 & 1.1059870 \\
      MCSA-ILUT-R-C & 1 & 25 & 1.1059870 \\
      MCSA-ILUT-MR-C & 1 & 25 & 1.1059870 \\
      MCSA-ILUT-R-EV & 1 & 25 & 1.1059870 \\
      MCSA-ILUT-MR-EV & 1 & 25 & 1.1059870 \\
      \hline
      BiCGStab-ILUT & 2 & 35 & 1.1552390 \\
      GMRES-ILUT & 2 & 35 & 1.1552390 \\
      MCSA-ILUT-R-C & 2 & 35 & 1.1552390 \\
      MCSA-ILUT-MR-C & 2 & 35 & 1.1552390 \\
      MCSA-ILUT-R-EV & 2 & 35 & 1.1552390 \\
      MCSA-ILUT-MR-EV & 2 & 35 & 1.1552390 \\
      \hline
      BiCGStab-ILUT & 4 & 31 & 1.1758500 \\
      GMRES-ILUT & 4 & 31 & 1.1758500 \\
      MCSA-ILUT-R-C & 4 & 31 & 1.1758500 \\
      MCSA-ILUT-MR-C & 4 & 31 & 1.1758500 \\
      MCSA-ILUT-R-EV & 4 & 31 & 1.1758500 \\
      MCSA-ILUT-MR-EV & 4 & 31 & 1.1758500 \\
      %%
      \hline\hline
    \end{tabular}
  \end{center}
  \caption{\textbf{Serial eigensolver verification results for the
      fuel assembly problem.} \textit{Each calculation was converged
      with the same eigensolver and linear solver tolerances.}}
  \label{tab:serial_ev_results}
\end{table}

@@

\begin{figure}[t!]
  \begin{center}
    \includegraphics[width=6in]{chapters/spn_equations/serial_1_group.pdf}
  \end{center}
  \caption{\textbf{Absolute difference of the group 0 scalar flux
      between the Aztec BiCGStab calculations and Aztec GMRES and the
      MCSA variations for the 1-group calculation.} \textit{Left to
      right: GMRES-ILUT, MCSA-ILUT-R-C, MCSA-ILUT-MR-C,
      MCSA-ILUT-R-EV, MCSA-ILUT-MR-EV.}}
  \label{fig:serial_1_group_diff}
\end{figure}

Table~\ref{tab:serial_differences_g1},
Table~\ref{tab:serial_differences_g2}, and
Table~\ref{tab:serial_differences_g4} give the maximum and minimum
absolute differences for all calculations and all energy groups using
the BiCGStab calculation as the reference. All comparisons were
performed within a group such that the MCSA-ILUT-R-C results in group
3 for the 4-group calculation were computed with the BiCGStab-ILUT
solution in group 3 for the 4-group calculation as a reference.

\begin{table}[h!]
  \begin{center}
    \begin{tabular}{lccc}\hline\hline
      \multicolumn{1}{c}{\textbf{Solver}} & 
      \multicolumn{1}{c}{\textbf{Group}} &
      \multicolumn{1}{c}{\textbf{Minimum Difference}} &
      \multicolumn{1}{c}{\textbf{Maximum Difference}} \\
      \hline
      GMRES-ILUT & 0 & 0.0 & \sn{9.313}{-10} \\
      MCSA-ILUT-R-C & 0 & 0.0 & \sn{9.313}{-10} \\
      MCSA-ILUT-MR-C & 0 & 0.0 & \sn{2.205}{-9} \\
      MCSA-ILUT-R-EV & 0 & 0.0 & \sn{9.313}{-10} \\
      MCSA-ILUT-MR-EV & 0 & 0.0 & \sn{9.313}{-10} \\
      \hline\hline
    \end{tabular}
  \end{center}
  \caption{\textbf{Serial scalar flux verification results for the
      1-group fuel assembly problem.} \textit{The maximum and minimum
      absolute differences over the entire spatial domain are
      presented with BiCGStab used as the reference. Each calculation
      was converged with the same eigensolver and linear solver
      tolerances.}}
  \label{tab:serial_differences_g1}
\end{table}

\begin{table}[h!]
  \begin{center}
    \begin{tabular}{lccc}\hline\hline
      \multicolumn{1}{c}{\textbf{Solver}} & 
      \multicolumn{1}{c}{\textbf{Group}} &
      \multicolumn{1}{c}{\textbf{Minimum Difference}} &
      \multicolumn{1}{c}{\textbf{Maximum Difference}} \\
      \hline
      GMRES-ILUT & 0 & 0.0 & \sn{2.794}{-9} \\
      MCSA-ILUT-R-C & 0 & 0.0 & \sn{9.313}{-10} \\
      MCSA-ILUT-MR-C & 0 & 0.0 & \sn{9.313}{-10} \\
      MCSA-ILUT-R-EV & 0 & 0.0 & \sn{9.313}{-10} \\
      MCSA-ILUT-MR-EV & 0 & 0.0 & \sn{9.313}{-10} \\
      \hline
      GMRES-ILUT & 1 & 0.0 & \sn{4.657}{-10} \\
      MCSA-ILUT-R-C & 1 & 0.0 & \sn{1.364}{-10} \\
      MCSA-ILUT-MR-C & 1 & 0.0 & \sn{1.261}{-9} \\
      MCSA-ILUT-R-EV & 1 & 0.0 & \sn{2.474}{-10} \\
      MCSA-ILUT-MR-EV & 1 & 0.0 & \sn{5.857}{-10} \\
      \hline\hline
    \end{tabular}
  \end{center}
  \caption{\textbf{Serial scalar flux verification results for the
      2-group fuel assembly problem.} \textit{The maximum and minimum
      absolute differences over the entire spatial domain are
      presented with BiCGStab used as the reference. Each calculation
      was converged with the same eigensolver and linear solver
      tolerances.}}
  \label{tab:serial_differences_g2}
\end{table}

\begin{table}[h!]
  \begin{center}
    \begin{tabular}{lccc}\hline\hline
      \multicolumn{1}{c}{\textbf{Solver}} & 
      \multicolumn{1}{c}{\textbf{Group}} &
      \multicolumn{1}{c}{\textbf{Minimum Difference}} &
      \multicolumn{1}{c}{\textbf{Maximum Difference}} \\
      \hline
      GMRES-ILUT & 0 & 0.0 & \sn{9.313}{-10} \\
      MCSA-ILUT-R-C & 0 & 0.0 & \sn{1.397}{-9} \\
      MCSA-ILUT-MR-C & 0 & 0.0 & \sn{1.870}{-9} \\
      MCSA-ILUT-R-EV & 0 & 0.0 & \sn{1.397}{-9} \\
      MCSA-ILUT-MR-EV & 0 & 0.0 & \sn{1.869}{-9} \\
      \hline
      GMRES-ILUT & 1 & 0.0 & \sn{5.821}{-11} \\
      MCSA-ILUT-R-C & 1 & 0.0 & \sn{1.164}{-10} \\
      MCSA-ILUT-MR-C & 1 & 0.0 & \sn{1.164}{-10} \\
      MCSA-ILUT-R-EV & 1 & 0.0 & \sn{1.164}{-10} \\
      MCSA-ILUT-MR-EV & 1 & 0.0 & \sn{1.164}{-10} \\
      \hline
      GMRES-ILUT & 2 & 0.0 & \sn{1.164}{-10} \\
      MCSA-ILUT-R-C & 2 & 0.0 & \sn{2.328}{-10} \\
      MCSA-ILUT-MR-C & 2 & 0.0 & \sn{2.328}{-10} \\
      MCSA-ILUT-R-EV & 2 & 0.0 & \sn{2.328}{-10} \\
      MCSA-ILUT-MR-EV & 2 & 0.0 & \sn{2.328}{-10} \\
      \hline
      GMRES-ILUT & 3 & 0.0 & \sn{1.164}{-10} \\
      MCSA-ILUT-R-C & 3 & 0.0 & \sn{1.746}{-10} \\
      MCSA-ILUT-MR-C & 3 & 0.0 & \sn{2.256}{-9} \\
      MCSA-ILUT-R-EV & 3 & 0.0 & \sn{2.801}{-10} \\
      MCSA-ILUT-MR-EV & 3 & 0.0 & \sn{2.256}{-10} \\
      \hline\hline
    \end{tabular}
  \end{center}
  \caption{\textbf{Serial scalar flux verification results for the
      4-group fuel assembly problem.} \textit{The maximum and minimum
      absolute differences over the entire spatial domain are
      presented with BiCGStab used as the reference. Each calculation
      was converged with the same eigensolver and linear solver
      tolerances.}}
  \label{tab:serial_differences_g4}
\end{table}

%%---------------------------------------------------------------------------%%
\section{MCSA Performance Comparison to Conventional Methods}
\label{sec:spn_comparison}
Using the same set of solver parameters used for verification, MCSA
will be directly compared to the production Krylov solvers in terms of
both iterative performance and CPU timing for solutions to the fuel
assembly problem. In addition to the parameters used for the
verification analysis, artificial absorption was also introduced with
a value of 0.2 for the MCSA calculations that used the collision
esitmator as CPU timing performance was improved without a loss in
iterative performance. Adding this absorption did not change the
numerical results of the calculation.

Figure~\ref{fig:spn_comparison_iterations} gives the average number of
linear solver iterations required to converge a single eigenvalue
iteration of the fuel assembly problem as a function of the number of
energy groups. In general, the iterative performance of all the
solvers tested is comparable with BiCGStab performing the best. We
expect this because not only are the multigroup $SP_N$ equations
elliptic and positive-definite, but they are also nearly symmetric and
therefore we expect the conjugate gradient-based methods to perform
well. It also important to note that the iterative performance of each
solver is not a strong function of the number of energy groups in the
problem (and therefore the problem size). For MCSA, we are linearly
increasing the number of stochastic histories used to compute the
correction at each iteration as the number of energy groups are
increased and therefore the work performed for the problem is of
$O(N)$. It is also important to note the the reduced domain
approximation parameters, particularly the fill level, were fixed as
the number of energy groups was increased. By fixing the fill level,
we are effectively fixing the amount of information contained in the
composite linear operator available for the Monte Carlo
problem. Limiting this information to 100 entries per row in the
preconditioned system did not appear to have a significant effect on
the iterative performance of the solver. For larger numbers of energy
groups (and therefore DOFs) this may be an issue and this number may
need to be increased to maintain iterative performance. Currently,
memory restrictions arising from the explicit preconditioning strategy
prevent larger energy groups to be analyzed using an appropriate
preconditioner for the fuel assembly problem.

@@

Although iterative performance for MCSA was comparable to that
observed for production Krylov implementations using exactly the same
preconditioning scheme, timing performance is not as
comparable. Figure~\ref{fig:spn_comparison_time} gives the average CPU
time per linear solver iteration required to converge the fuel
assembly problem with all linear solver iterations over all eigenvalue
iterations used to compute the average. The production Krylov solver
implementations are approximately an order of magnitude faster than
the MCSA implementations presented here. We expect these results for
several reasons. First, even when the reduced domain approximation is
used there are $O(100)$ elements in each row of the system and each of
those elements must be processed during a transition in the Monte
Carlo random walk sequence. In addition, although the preconditioned
fixed point iterations used in the MCSA sequence do not use the
composite linear operator but rather a sequence of matrix vector
multiplies to achieve the same preconditioning effect, they do use the
explicitly extracted inverse of the preconditioning matrices which
themselves are dense, leading to exceedingly slow computation times
even in the fixed point iteration. For the production Krylov methods,
the ILUT preconditioners are applied to a vector with two triangular
solves, one for each triangular factor. Also, the original linear
operator has $O(10)$ elements in each row of the system, an order of
magnitude less than that used in the reduced domain composite
operator. Second, the implementation presented here has not been
optimized and it is likely that several components of the algorithm
can be implemented in a more desireable time complexity. The
production Krylov solvers presented here have enjoyed decades of
professional development and optimization. However, it is most
important to note here that the CPU timing data presented in
Figure~\ref{fig:spn_comparison_time} shows that all methods have the
same time complexity; their differences are merely the manifestation
of a large time constant due to the effects of the explicit
preconditioning scheme. For transport systems where this is not a
problem, the literature explicitly shows that MCSA can be competetive
with these production Krylov methods using CPU time as a measure
\citep{evans_monte_2012}.

>>>>

For this study, the fuel assemby problem with a 1-group $SP_1$
discretization was solved using MCSA with the adjoint Monte Carlo
solver using both the collision and expected value estimators and both
the Richardson and MINRES fixed point iterations. Using the results
from the previous section, a Neumann relaxation parameter of 0.7 was
used with both estimators and fixed point iterations. For the
collision estimator, a Richardson relaxation parameter of 1.1 was used
with the Richardson iteration while it was found the expected value
estimator had the best performance when a Richardson relaxation
parameter of 1.0 was used instead. In addition, ILUT preconditioning
with a drop tolerance of \sn{1}{-5} and a fill level of 5 was used
with a reduced domain approximation fill level of 100 and a threshold
of \sn{1}{-10} as in the previous analysis. Artificial absorption was
also introduced with a value of 0.2 for the MCSA calculations that
used the collision estimator as CPU timing performance was improved
without a loss in iterative performance while no absorption was used
for calculations using the expected value estimator. For each
estimator, the number of stochastic histories used to compute the MCSA
correction was varied from 25 to 50,000 with the number of MCSA
iterations required to converge to a tolerance of \sn{1}{-8} recorded
for a single eigenvalue iteration along with the CPU time required to
achieve convergence.

@@

%%---------------------------------------------------------------------------%%
\section{MCSA Verification}
\label{sec:spn_mcsa_verification}
Through the numerical studies presented in this chapter, we have
developed a set of preconditioning techniques that permit MCSA to be
used with the $SP_N$ form of the neutron transport equation and
applied to a difficult fuel assembly criticality problem. Along with
these preconditioning techniques, a set a varying MCSA parameters were
analyzed to find those that yielded the best performance for this
particular problem. In addition, several steps were taken to mitigate
the dense composite linear operators that arise when performing
explicit MCSA preconditioning in order to achieve convergence. Using
these results, we compare MCSA directly to conventional linear solvers
that would be typically used to solve the $SP_N$ form of the transport
equation as presented here. In particular, we will use a pair of
production Krylov solvers with which we will compare direct numerical
results in order to verify MCSA solutions in this section.

@@

For every combination of energy group structure and solver, the
eigenvalue problem was converged with an eigensolver tolerance of
\sn{1}{-6} and a linear solver tolerance of \sn{1}{-8} in a serial
calculation. Table~\ref{tab:serial_ev_results} gives the eigenvalue
computed by each solver for each problem as well as the number of
eigenvalue iterations required to converge. For every problem, each
solver converged to the same eigenvalue within the tolerance of the
eigensolver in the same number of eigenvalue iterations.
\begin{table}[h!]
  \begin{center}
    \begin{tabular}{lccc}\hline\hline
      \multicolumn{1}{c}{\textbf{Solver}} & 
      \multicolumn{1}{c}{\textbf{Number of Groups}} & 
      \multicolumn{1}{c}{\textbf{Iterations}} &
      \multicolumn{1}{c}{\textbf{Eigenvalue}} \\
      \hline
      BiCGStab-ILUT & 1 & 25 & 1.1059870 \\
      GMRES-ILUT & 1 & 25 & 1.1059870 \\
      MCSA-ILUT-R-C & 1 & 25 & 1.1059870 \\
      MCSA-ILUT-MR-C & 1 & 25 & 1.1059870 \\
      MCSA-ILUT-R-EV & 1 & 25 & 1.1059870 \\
      MCSA-ILUT-MR-EV & 1 & 25 & 1.1059870 \\
      \hline
      BiCGStab-ILUT & 2 & 35 & 1.1552390 \\
      GMRES-ILUT & 2 & 35 & 1.1552390 \\
      MCSA-ILUT-R-C & 2 & 35 & 1.1552390 \\
      MCSA-ILUT-MR-C & 2 & 35 & 1.1552390 \\
      MCSA-ILUT-R-EV & 2 & 35 & 1.1552390 \\
      MCSA-ILUT-MR-EV & 2 & 35 & 1.1552390 \\
      \hline
      BiCGStab-ILUT & 4 & 31 & 1.1758500 \\
      GMRES-ILUT & 4 & 31 & 1.1758500 \\
      MCSA-ILUT-R-C & 4 & 31 & 1.1758500 \\
      MCSA-ILUT-MR-C & 4 & 31 & 1.1758500 \\
      MCSA-ILUT-R-EV & 4 & 31 & 1.1758500 \\
      MCSA-ILUT-MR-EV & 4 & 31 & 1.1758500 \\
      %%
      \hline\hline
    \end{tabular}
  \end{center}
  \caption{\textbf{Serial eigensolver verification results for the
      fuel assembly problem.} \textit{Each calculation was converged
      with the same eigensolver and linear solver
      tolerances. Table~\ref{tab:spn_solver_defs} gives the
      description for each solver type presented.}}
  \label{tab:serial_ev_results}
\end{table}

@@

\begin{figure}[t!]
  \begin{center}
    \includegraphics[width=6in]{chapters/spn_equations/serial_1_group.pdf}
  \end{center}
  \caption{\textbf{Absolute difference of the group 0 scalar flux
      between the Aztec BiCGStab calculations and Aztec GMRES and the
      MCSA variations for the 1-group calculation.} \textit{Left to
      right: GMRES-ILUT, MCSA-ILUT-R-C, MCSA-ILUT-MR-C,
      MCSA-ILUT-R-EV, MCSA-ILUT-MR-EV. Table~\ref{tab:spn_solver_defs}
      gives the description for each solver type presented.}}
  \label{fig:serial_1_group_diff}
\end{figure}

Table~\ref{tab:serial_differences_g1},
Table~\ref{tab:serial_differences_g2}, and
Table~\ref{tab:serial_differences_g4} give the maximum and minimum
absolute differences for all calculations and all energy groups using
the BiCGStab calculation as the reference. All comparisons were
performed within a group such that the MCSA-ILUT-R-C results in group
3 for the 4-group calculation were computed with the BiCGStab-ILUT
solution in group 3 for the 4-group calculation as a reference. For
all group flux results given in the table, the maximum differences for
all cases was less than the linear solver tolerance used to converge
the flux at each eigenvalue iteration meaning that all values over the
entire energy and spatial domain are within these limits. In addition,
the minimum error for all calculations was observed to be zero,
meaning in at least one mesh cell in all calculations for all energy
groups the scalar flux values computed by all solvers were
identical. Based on the flux results and the eigenvalue results
presented here, the MCSA results should be deemed correct for the fuel
assembly problem for serial calculations.

\begin{table}[h!]
  \begin{center}
    \begin{tabular}{lccc}\hline\hline
      \multicolumn{1}{c}{\textbf{Solver}} & 
      \multicolumn{1}{c}{\textbf{Group}} &
      \multicolumn{1}{c}{\textbf{Minimum Difference}} &
      \multicolumn{1}{c}{\textbf{Maximum Difference}} \\
      \hline
      GMRES-ILUT & 0 & 0.0 & \sn{9.313}{-10} \\
      MCSA-ILUT-R-C & 0 & 0.0 & \sn{9.313}{-10} \\
      MCSA-ILUT-MR-C & 0 & 0.0 & \sn{2.205}{-9} \\
      MCSA-ILUT-R-EV & 0 & 0.0 & \sn{9.313}{-10} \\
      MCSA-ILUT-MR-EV & 0 & 0.0 & \sn{9.313}{-10} \\
      \hline\hline
    \end{tabular}
  \end{center}
  \caption{\textbf{Serial scalar flux verification results for the
      1-group fuel assembly problem.} \textit{The maximum and minimum
      absolute differences over the entire spatial domain are
      presented with BiCGStab used as the reference. Each calculation
      was converged with the same eigensolver and linear solver
      tolerances. Table~\ref{tab:spn_solver_defs} gives the
      description for each solver type presented.}}
  \label{tab:serial_differences_g1}
\end{table}

\begin{table}[h!]
  \begin{center}
    \begin{tabular}{lccc}\hline\hline
      \multicolumn{1}{c}{\textbf{Solver}} & 
      \multicolumn{1}{c}{\textbf{Group}} &
      \multicolumn{1}{c}{\textbf{Minimum Difference}} &
      \multicolumn{1}{c}{\textbf{Maximum Difference}} \\
      \hline
      GMRES-ILUT & 0 & 0.0 & \sn{2.794}{-9} \\
      MCSA-ILUT-R-C & 0 & 0.0 & \sn{9.313}{-10} \\
      MCSA-ILUT-MR-C & 0 & 0.0 & \sn{9.313}{-10} \\
      MCSA-ILUT-R-EV & 0 & 0.0 & \sn{9.313}{-10} \\
      MCSA-ILUT-MR-EV & 0 & 0.0 & \sn{9.313}{-10} \\
      \hline
      GMRES-ILUT & 1 & 0.0 & \sn{4.657}{-10} \\
      MCSA-ILUT-R-C & 1 & 0.0 & \sn{1.364}{-10} \\
      MCSA-ILUT-MR-C & 1 & 0.0 & \sn{1.261}{-9} \\
      MCSA-ILUT-R-EV & 1 & 0.0 & \sn{2.474}{-10} \\
      MCSA-ILUT-MR-EV & 1 & 0.0 & \sn{5.857}{-10} \\
      \hline\hline
    \end{tabular}
  \end{center}
  \caption{\textbf{Serial scalar flux verification results for the
      2-group fuel assembly problem.} \textit{The maximum and minimum
      absolute differences over the entire spatial domain are
      presented with BiCGStab used as the reference. Each calculation
      was converged with the same eigensolver and linear solver
      tolerances. Table~\ref{tab:spn_solver_defs} gives the
      description for each solver type presented.}}
  \label{tab:serial_differences_g2}
\end{table}

\begin{table}[h!]
  \begin{center}
    \begin{tabular}{lccc}\hline\hline
      \multicolumn{1}{c}{\textbf{Solver}} & 
      \multicolumn{1}{c}{\textbf{Group}} &
      \multicolumn{1}{c}{\textbf{Minimum Difference}} &
      \multicolumn{1}{c}{\textbf{Maximum Difference}} \\
      \hline
      GMRES-ILUT & 0 & 0.0 & \sn{9.313}{-10} \\
      MCSA-ILUT-R-C & 0 & 0.0 & \sn{1.397}{-9} \\
      MCSA-ILUT-MR-C & 0 & 0.0 & \sn{1.870}{-9} \\
      MCSA-ILUT-R-EV & 0 & 0.0 & \sn{1.397}{-9} \\
      MCSA-ILUT-MR-EV & 0 & 0.0 & \sn{1.869}{-9} \\
      \hline
      GMRES-ILUT & 1 & 0.0 & \sn{5.821}{-11} \\
      MCSA-ILUT-R-C & 1 & 0.0 & \sn{1.164}{-10} \\
      MCSA-ILUT-MR-C & 1 & 0.0 & \sn{1.164}{-10} \\
      MCSA-ILUT-R-EV & 1 & 0.0 & \sn{1.164}{-10} \\
      MCSA-ILUT-MR-EV & 1 & 0.0 & \sn{1.164}{-10} \\
      \hline
      GMRES-ILUT & 2 & 0.0 & \sn{1.164}{-10} \\
      MCSA-ILUT-R-C & 2 & 0.0 & \sn{2.328}{-10} \\
      MCSA-ILUT-MR-C & 2 & 0.0 & \sn{2.328}{-10} \\
      MCSA-ILUT-R-EV & 2 & 0.0 & \sn{2.328}{-10} \\
      MCSA-ILUT-MR-EV & 2 & 0.0 & \sn{2.328}{-10} \\
      \hline
      GMRES-ILUT & 3 & 0.0 & \sn{1.164}{-10} \\
      MCSA-ILUT-R-C & 3 & 0.0 & \sn{1.746}{-10} \\
      MCSA-ILUT-MR-C & 3 & 0.0 & \sn{2.256}{-9} \\
      MCSA-ILUT-R-EV & 3 & 0.0 & \sn{2.801}{-10} \\
      MCSA-ILUT-MR-EV & 3 & 0.0 & \sn{2.256}{-10} \\
      \hline\hline
    \end{tabular}
  \end{center}
  \caption{\textbf{Serial scalar flux verification results for the
      4-group fuel assembly problem.} \textit{The maximum and minimum
      absolute differences over the entire spatial domain are
      presented with BiCGStab used as the reference. Each calculation
      was converged with the same eigensolver and linear solver
      tolerances. Table~\ref{tab:spn_solver_defs} gives the
      description for each solver type presented.}}
  \label{tab:serial_differences_g4}
\end{table}

%%---------------------------------------------------------------------------%%
\section{MCSA Performance Comparison to Conventional Methods}
\label{sec:spn_comparison}
In the previous section, numerical results using MCSA for the fuel
assembly calculation were directly with production Krylov methods in
order to verify their correctness. Deemed correct, with the same set
of solver parameters used for verification MCSA will now be directly
compared to the same production Krylov solvers in terms of both
iterative performance and CPU timing for solutions to the fuel
assembly problem.  This comparison of both iterative and CPU timing
performance is necessary in order to gauge feasibility of the method
and identify areas that require further research and development.

To compare iterative performance, the number of linear solver
iterations required to converge each eigenvalue iteration was recorded
and then averaged over all eigenvalue iterations for each
solver. Figure~\ref{fig:spn_comparison_iterations} gives the average
number of linear solver iterations required to converge a single
eigenvalue iteration of the fuel assembly problem as a function of the
number of energy groups. In general, the iterative performance of all
the solvers tested is comparable with BiCGStab performing the best. We
expect this because not only are the multigroup $SP_N$ equations
elliptic and positive-definite, but they are also nearly symmetric and
therefore we expect the conjugate gradient-based methods to perform
well. It also important to note that the iterative performance of each
solver is not a strong function of the number of energy groups in the
problem (and therefore the problem size). For MCSA, we are linearly
increasing the number of stochastic histories used to compute the
correction at each iteration as the number of energy groups are
increased and therefore the work performed for the problem is of
$O(N)$. It is also important to note the reduced domain approximation
parameters, particularly the fill level, were fixed as the number of
energy groups was increased. By fixing the fill level, we are
effectively fixing the amount of information contained in the
composite linear operator available for the Monte Carlo
problem. Limiting this information to 100 entries per row in the
preconditioned system did not appear to have a significant effect on
the iterative performance of the solver. For larger numbers of energy
groups (and therefore DOFs) this may be an issue and this number may
need to be increased to maintain iterative performance. Currently,
memory restrictions arising from the explicit preconditioning strategy
prevent larger energy groups to be analyzed using an appropriate
preconditioner for the fuel assembly problem.

@@

Although iterative performance for MCSA was comparable to that
observed for production Krylov implementations using exactly the same
preconditioning scheme, timing performance is not as competetive. For
this comparison, the amount of time to perform each linear solver
iteration in every eigenvalue iteration was recorded and then averaged
over all linear solver
iterations. Figure~\ref{fig:spn_comparison_time} gives the average CPU
time per linear solver iteration required to converge the fuel
assembly problem with all linear solver iterations over all eigenvalue
iterations used to compute the average. The production Krylov solver
implementations are approximately an order of magnitude faster than
the MCSA implementations presented here. We expect these results for
several reasons. First, even when the reduced domain approximation is
used there are $O(100)$ elements in each row of the system and each of
those elements must be processed during a transition in the Monte
Carlo random walk sequence. In addition, although the preconditioned
fixed point iterations used in the MCSA sequence do not use the
composite linear operator but rather a sequence of matrix vector
multiplies to achieve the same preconditioning effect, they do use the
explicitly extracted inverse of the preconditioning matrices which
themselves are dense, leading to exceedingly slow computation times
even in the fixed point iteration. For the production Krylov methods,
the ILUT preconditioners are applied to a vector with two triangular
solves, one for each triangular factor. Also, the original linear
operator has $O(10)$ elements in each row of the system, an order of
magnitude less than that used in the reduced domain composite
operator. Second, the implementation presented here has not been
optimized and it is likely that several components of the algorithm
can be implemented in a more desireable time complexity. The
production Krylov solvers presented here have enjoyed decades of
professional development and optimization. However, it is most
important to note here that the CPU timing data presented in
Figure~\ref{fig:spn_comparison_time} shows that all methods have the
same time complexity; their differences are merely the manifestation
of a large time constant due to the effects of the explicit
preconditioning scheme. For transport systems where this is not a
problem, the literature explicitly shows that MCSA can be competetive
with these production Krylov methods using CPU time as a measure
\citep{evans_monte_2012}.
